Concurrent Programming in Harmony

### 2021-12-03 book/paper.tex
Starting to incorporate nice pics into the book
Starting to work on book again...

So, what is Harmony?
Harmony is a concurrent programming language.  It was designed to teach
the basics of concurrent and distributed programming, but it is also useful for
testing new concurrent algorithms or even sequential and distributed
algorithms.  Harmony programs are not intended to be ``run'' like programs
in most other programming languages---instead Harmony programs are
model checked to test that the program has certain desirable
properties and does not suffer from bugs.
Interestingly, Harmony does not have input/output statements, so there is
no ``Hello World'' program.

@@

>>>>

So, what is Harmony?
Harmony is a concurrent programming language.  It was designed to teach
the basics of concurrent and distributed programming, but it is also useful for
testing new concurrent algorithms or even sequential and distributed
algorithms.  Harmony programs are not intended to be ``run'' like programs
in most other programming languages---instead Harmony programs are
model checked to test that the program has certain desirable
properties and does not suffer from bugs.

@@

\chapter{Hello World}
\label{ch:helloworld}

\begin{figure}[h]
\begin{center}
\begin{tabular}{ccc}
\includegraphics[width=0.35\textwidth]{../code/hello1.png}
& \rule{0.05\linewidth}{0in} &
\includegraphics[width=0.45\textwidth]{../code/hello2.png} \\
\begin{tcolorbox}[width=0.4\linewidth]
\harmonysource{hello1}
\end{tcolorbox}
& \rule{0.05\linewidth}{0in} &
\begin{tcolorbox}[width=0.4\linewidth]
\harmonysource{hello2}
\end{tcolorbox}
\\
(a) \harmonylink{code/hello1.hny} && (b) \harmonylink{code/hello2.hny}
\end{tabular}
\end{center}
\caption{Hello World!}
\label{fig:helloworld}
\end{figure}

The first programming book I read cover to cover was \emph{The C Programming
Language} (first edition) by Brian W. Kernighan and Dennis M. Ritchie, which
was around 1980.  In 1990, Dennis was my boss at AT\&T Bell Labs, while Brian
was my colleague in the same lab.  The first program in the book printed
the string ``hello, world''.  Since then, most programming tutorials for
pretty much any programming language start with that example.  Harmony, too,
has a Hello World program.
\autoref{fig:helloworld}(a) shows the program and the corresponding output.
The output is a \emph{Deterministic State Machine} (DFA).  The green circle
represents the initial state and the double circle represents the final
state.  There is one \emph{transition}, labeled with the string
``hello world''.  The DFA describes (or \emph{recognizes}) all possible
outputs that the program can generate.  In this case, there is only one.

Now look at \autoref{fig:helloworld}{b).  It has two \texttt{print} statements.
The output is another DFA, although this time there is an intermediate state.
Still, there is only one possible output, and it consists of the strings
``hello'' and ``world''.

But programs can usually have more than one execution and produce multiple
different outputs as a result.  \autoref{fig:choice} demonstrates
\emph{nondetermistic choice} in Harmony programs.  In this case, the
program chooses to either print ``hello'' or ``world''.  The corresponding
DFA captures both possibilities.

\begin{figure}
\begin{center}
\includegraphics[width=0.45\textwidth]{../code/hello3.png}
\end{center}
\begin{code}
\harmonysource{hello3}
\end{code}
\caption{\harmonylink{code/hello3.hny} Nondeterministic choice}
\label{fig:choice}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.15\textwidth]{../code/hello4.png}
\end{center}
\begin{code}
\harmonysource{hello4}
\end{code}
\caption{\harmonylink{code/hello4.hny} Loop for some time, but don't have to
go on for ever}
\label{fig:hello4}
\end{figure}

\begin{figure}[h]
\begin{center}
\begin{tabular}{ccc}
\includegraphics[width=0.45\textwidth]{../code/hello5.png}
& \rule{0.05\linewidth}{0in} &
\includegraphics[width=0.45\textwidth]{../code/hello6.png} \\
\begin{tcolorbox}[width=0.4\linewidth]
\harmonysource{hello5}
\end{tcolorbox}
& \rule{0.05\linewidth}{0in} &
\begin{tcolorbox}[width=0.4\linewidth]
\harmonysource{hello6}
\end{tcolorbox}
\\
(a) \harmonylink{code/hello5.hny} && (b) \harmonylink{code/hello6.hny}
\end{tabular}
\end{center}
\caption{Method calls and threads}
\label{fig:methodthread}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.9\textwidth]{../code/hello7.png}
\end{center}
\begin{code}
\harmonysource{hello7}
\end{code}
\caption{\harmonylink{code/hello7.hny} Interleaving threads}
\label{fig:hello7}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{../code/hello8.png}
\end{center}
\begin{code}
\harmonysource{hello8}
\end{code}
\caption{\harmonylink{code/hello8.hny} More interleaving}
\label{fig:hello8}
\end{figure}
